Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2020-05-14 paper2_JNLS/Referee Reports/Ref2Response/Ref2.tex
NM

\noindent
We have replaced the majority of these descriptors with more quantitative ones. In particular:
\begin{itemize}
\item The term `spectacular fashion' in the abstract was changed to `convincing evidence'.
\item The word `modest' on page 18 was replaced with `moderate'. Thanks for pointing this out -- I believe `moderate' was the word that was originally intended and `modest' was the result of an autocorrect.
\item The word `roughly' (appearing in several places in the original manuscript) was either replaced by `approximately' or else removed.
\item The phrase `captures the trend remarkably well' on page 20 was changed to `captures the trend closely'.
\item We have kept `intriguing' on page 17 as we believe it to be an appropriate use of the word.
\end{itemize}
\hphantom \\

@@

\begin{itemize}
\item As can bee seen in Fig. 2 (b) and (c), the location of the peak is indeed insensitive to driving amplitude to within experimental error (the level of which can be estimated by the noise in the data). Accordingly, we have replaced `appears insensitive' to simply `is insensitive'.
We had initially used the word `appears' to encourage the reader to view the figure and see for themselves. However, we now see the referee's point that this word may introduce some uncertainty in the reader's mind and we have thus made the change.
\end{itemize}

@@

\begin{itemize}
\item This is a great question and one that we have already devoted some energy to understanding. First, it is unclear whether the transfer function converges to a single curve in the limit $\Lambda \to \infty$ and this is a question currently under investigation (see more below). Second, to briefly explain our choices, we selected the system size $\Lambda$ as a moderate value that is large enough to capture sufficiently rich dynamics and small enough to be computationally tractable. Meanwhile, we chose $N = \Lambda / 2$ so that the dominant forcing frequency $f_p$ sits near the middle of resolved scales (logarithmically) in the model. See sections 3.4 and especially 3.9 in the manuscript for a detailed discussion of the parameter $N$. In particular, Section 3.9 outlines a simple scaling argument that suggests scaling $N$ linearly with $\Lambda$ offsets the sensitivity of the transfer function to $\Lambda$ for asymptotically large $\Lambda$ (it is unclear, however, if this mitigates the sensitivity enough).

@@

% COMMENT 6
\noindent
\comment{Sharp steps do occur in bathymetry profiles. However, do you think Eq (53) will still apply rather well over other bathymetry profiles such as a transition with a steep finite gradient?}

\begin{itemize}
\item
\end{itemize}

% COMMENT 7
\noindent
\comment{Minor typos etc.
Pg 6 Ln 60: is leading-order -> is the leading-order
Pg 7 Ln 21: change dash to en or em dash as appropriate
Pg 16 Ln 35: Fig. 3(b) -> Fig. 3(c)
Pg 24: References [17] and [24] require capitals as they are book titles, and [27] requires a capital H for Hamiltonian}

\begin{itemize}
\item
\end{itemize}

% COMMENT 8
\noindent
\comment{}

\begin{itemize}
\item
\end{itemize}
\end{document}

>>>>

\noindent
We have replaced the majority of these descriptors with more quantitative ones. In particular:
\begin{itemize}
\item The term `spectacular fashion' in the abstract was changed to `convincing evidence'.
\item The word `modest' on page 18 was replaced with `moderate'. Thanks for pointing this out -- I believe `moderate' was the word that was originally intended and `modest' was the result of an autocorrect.
\item The word `roughly' (appearing in several places in the original manuscript) was either replaced by `approximately' or else removed.
\item The phrase `captures the trend remarkably well' on page 20 was changed to `the data is well predicted by formula (53)'.
\item We have kept `intriguing' on page 17 as we believe it to be an appropriate use of the word.
\end{itemize}
\hphantom \\

@@

\begin{itemize}
\item As can bee seen in Fig. 2 (b) and (c), the location of the peak is indeed insensitive to driving amplitude to within experimental error (the level of error can be visually estimated by the noise in the data). Accordingly, we have replaced `appears insensitive' to simply `is insensitive'.
We had initially used the word `appears' to encourage the reader to view the figure and see for themselves. However, we now see the referee's point that this word may weaken our statement and introduce some uncertainty in the reader's mind.
\end{itemize}

@@

\begin{itemize}
\item This is a great question and one that we have already devoted some energy to understanding. First, it is unclear whether the transfer function converges to a single curve in the limit $\Lambda \to \infty$ and this is a question currently under investigation (see more below). Second, to briefly explain our choices, we selected the system size $\Lambda$ as a moderate value that is large enough to capture sufficiently rich dynamics and small enough to be computationally tractable. Meanwhile, we chose $N = \Lambda / 2$ so that the dominant forcing frequency $f_p$ sits near the middle of resolved scales (logarithmically) in the model. See sections 3.4 and especially 3.9 in the manuscript for a detailed discussion of the parameter $N$. In particular, Section 3.9 outlines a simple scaling argument that suggests scaling $N$ linearly with $\Lambda$ offsets the sensitivity of the transfer function to $\Lambda$ for asymptotically large $\Lambda$ (it is unclear, however, if the sensitivity has been mitigated enough).

@@

% COMMENT 6
\noindent
\comment{Sharp steps do occur in bathymetry profiles. However, do you think Eq (53) will still apply rather well over other bathymetry profiles such as a transition with a steep finite gradient?}
\begin{itemize}
\item This is a good question. Since the wave response depends continuously on bottom topography, one would expect a surface of finite, steep slope to produce behavior similar to what we observe. How steep the surface must be for a reasonable comparison is a more difficult question, but perhaps some insight can be gained through back-of-the-envelope scaling analysis. The relevant length-scale for the wave response is the wavelength. Essentially, waves cannot ``see'' scales that are much smaller. Hence, a bathymetric change that takes place over a horizontal scale much smaller than a wavelength would be nearly indistinguishable from an abrupt depth change. Upon reflection, this line of thought implies that it is not necessarily the slope of the surface that matters most. For example, if the water depth is much smaller than the wavelength, then a mild slope could cause an appreciable change in depth that takes place over a horizontal distance much smaller than the wavelength. To give a concrete (yet somewhat exaggerated) example, suppose depth/wavelength = 1/10 and the bottom slope is 1/5. Then a change of 50\% in the depth occurs over 1/4 of a wavelength. Hence, in this scenario, a mild bottom slope of 1/5 might reasonably be approximated by an abrupt depth change.
\end{itemize}

% COMMENT 7
\noindent
\comment{Minor typos etc.
\\Pg 6 Ln 60: is leading-order $\to$ is the leading-order %Corrected
\\Pg 7 Ln 21: change dash to en or em dash as appropriate % Corrected
\\Pg 16 Ln 35: Fig. 3(b) $\to$ Fig. 3(c)%Corrected
\\Pg 24: References [17] and [24] require capitals as they are book titles, and [27] requires a capital H for Hamiltonian %Corrected
}
\begin{itemize}
\item We have corrected these typos. Thank you for bringing them to our attention.
\end{itemize}

\end{document}

### 2020-05-14 paper2_JNLS/paper.tex
NM

We consider the surface displacement $\eta(x,t)$ of unidirectional, shallow-water waves in a reference frame moving with the characteristic wave speed, $\xi = x - ct$. Here, $c = \sqrt{g \depth}$ is {the} leading-order approximation to the wave speed (i.e.~from linear theory), where $g$ is gravity and $\depth$ is the local depth.
The leading-order dynamics are corrected to first order in small amplitude by the Korteweg–de Vries equation (KdV), which in dimensional form is given by \cite{whitham2011linear}
\begin{equation}
\label{KdV}
\eta_t + \frac{3 c}{2 \depth} \eta \eta_{\xi} + \frac{c \depth^2}{6} \eta_{\xi \xi \xi} = 0
\end{equation}
% Given by Whitham on p. 461, Eq. 13.98
Motivated by the experiments, we consider waves that originate from a region of constant depth, encounter an abrupt depth change, and continue into another region of constant depth. Thus, depth will be piecewise constant
\begin{align}
\depth = 
\begin{cases}
\dup \quad \mbox{if } x<0 \\
\ddn \quad \mbox{if } x>0
\end{cases}
\end{align}
Most often, we consider waves moving into shallower depth, so that $\dup > \ddn$. Throughout this paper, we use the subscript `-' to represent upstream variables and `+' for downstream variables.

@@

\new{We comment on some important differences between the KdV and TKdV dynamical systems. Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants. Trajectories of KdV are therefore constrained by an infinite number of conserved quantities, associated with the fact that it is a completely integrable system whose long-time dynamics are dominated by coherent structures known as solitons. The analogues of these additional invariants, however, are {\em not} generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian. Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous mathematical proof awaits) \cite{bajars2013weakly}. In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, where  integrability is exploited to uncover certain statistical features. We remind the reader that KdV is asymptotically valid only for large spatial scales and hence truncating the small scales does not alter the order of approximation made. In this sense, KdV and TKdV should be regarded as on equal footing as far as physical realism, provided that sufficiently many modes are considered. Where differences exist in long-time dynamics, it is unclear {\em a priori} which model, if either, more closely resembles reality, which underscores the importance of controlled laboratory experiments.
}

@@

	We note that, in the current formulation, a positive inverse temperature $\invtemp > 0$ produces physically realistic statistics with a decaying energy spectrum, as is consistent with experiments \cite{majda2019, bajars2013weakly}. Negative inverse temperature produces a physically unrealistic spectrum that has more energy at smaller scales. Hence, we will hereafter focus on the physically realistic case of $\invtemp > 0$. \new{We also point out that a sufficient amount of nonlinear mixing is required for relaxation to the above Gibbs measure. In the linearized system, for example, modes propagate without interacting and hence the initial measure induced by the wave generator remains unchanged.}

%-------------------- COMMENTED OUT ------------------------%
% Definitions of ensemble and time averages.
% It seems that I did not need to use this notation so lets kill it.
\begin{comment}
Throughout this paper we use a bracket for ensemble average and overbar for a time average; i.e.~for any quantity $Q$,
\begin{align}
&\mean{Q} = \int_{\Omega}  Q \, d \Gibbs
&&\mbox{\em Ensemble average} \\
&\tavg{Q} = \frac{1}{T_2-T_1} \int_{T_1}^{T_2} Q \, dt	 
&&\mbox{\em Time average}
\end{align}
Note that if $\invtemp = 0$, then the Gibbs measure is a {\em uniform} measure on the unit hypersphere in $\CC^{\Lambda}$, which we will denote as either $\Gz$ or simply $d\mu$.
\end{comment}
%----------------------------------------------------------------------------%

>>>>

We consider the surface displacement $\eta(x,t)$ of unidirectional, shallow-water waves in a reference frame moving with the characteristic wave speed, $\xi = x - ct$. Here, $c = \sqrt{g \depth}$ is {the} leading-order approximation to the wave speed (i.e.~from linear theory), where $g$ is gravity and $\depth$ is the local depth.
The leading-order dynamics are corrected to first order in small amplitude by the Korteweg–de Vries equation (KdV), which in dimensional form is given by \cite{whitham2011linear}
\begin{equation}
\label{KdV}
\eta_t + \frac{3 c}{2 \depth} \eta \eta_{\xi} + \frac{c \depth^2}{6} \eta_{\xi \xi \xi} = 0
\end{equation}
% Given by Whitham on p. 461, Eq. 13.98
Motivated by the experiments, we consider waves that originate from a region of constant depth, encounter an abrupt depth change, and continue into another region of constant depth. Thus, depth will be piecewise constant
\begin{align}
\depth = 
\begin{cases}
\dup \quad \mbox{if } x<0 \\
\ddn \quad \mbox{if } x>0
\end{cases}
\end{align}
Most often, we consider waves moving into shallower depth, so that $\dup > \ddn$. Throughout this paper, we use the subscript `$-$' and to represent upstream variables and `$+$' for downstream variables.

@@

\new{We comment on some important differences between the KdV and TKdV dynamical systems. Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants. Trajectories of KdV are therefore constrained by an infinite number of conserved quantities, associated with the fact that it is a completely integrable system whose long-time dynamics are dominated by coherent structures known as solitons. The analogues of these additional invariants, however, are {\em not} generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian. Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous mathematical proof awaits) \cite{bajars2013weakly}. In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, where  integrability is exploited to uncover certain statistical features. We remind the reader that KdV is asymptotically valid only for large spatial scales and hence truncating the small scales does not alter the order of approximation made. In this sense, KdV and TKdV should be regarded as on equal footing as far as physical realism, provided that sufficiently many modes are considered. Where differences exist in long-time dynamics, it is unclear {\em a priori} which model, if either, more closely resembles reality, underscoring the importance of controlled laboratory experiments.
}

@@

	We note that, in the current formulation, a positive inverse temperature $\invtemp > 0$ produces physically realistic statistics with a decaying energy spectrum, as is consistent with experiments \cite{majda2019, bajars2013weakly}. Negative inverse temperature produces a physically unrealistic spectrum that has more energy at smaller scales. Hence, we will hereafter focus on the physically realistic case of $\invtemp > 0$. \new{We also point out that a sufficient amount of nonlinear mixing is required for relaxation to the above Gibbs measure. In the linearized system, for example, modes propagate without interacting, and hence the initial measure induced by the wave generator would remain unchanged.}
